---
title: "data_manipulation"
author: "Shihui Peng"
date: "2023-09-28"
output: github_document
---

```{r, echo=FALSE, message=FALSE}
library(tidyverse)
```

Import FAS litters and pups.

```{r}
litters_df = 
  read_csv("data/FAS_litters.csv")

litters_df = 
  janitor::clean_names(litters_df)

pups_df = 
  read_csv("data/FAS_pups.csv")

pups_df = 
  janitor::clean_names(pups_df)
```

# `select`

```{r}
select(litters_df, group, litter_number, gd0_weight)

select(litters_df, gd0_weight, litter_number, group)

select(litters_df, group, gd0_weight:gd_of_birth)

select(litters_df, group, starts_with("pups"))
```

* **`select`**: is used to select columns. If want to extract only some columns, we can use this.
when using it: **`select(data_frame_Im_working_on_now, column1_want_to_keep, column2_want_to_keep, ...)`**. If put in diff order, r will put these columns in the order we have them in. So we can reorganize data in this way.

* select(litters_df, group, **gd0_weight : gd_of_birth**): keep group col & "A:B" here means we want everything between A and B

* select(litters_df, group, **starts_with("pups")**): keep group col & "starts_with("ABC")" here means select everything that starts with ABC. There are also **`ends_with`** and **`contains`**

## `select` and remove cols while keeping all other cols
```{r}
select(litters_df, - litter_number)

select(litters_df, - starts_with("gd"))
```

* select(litters_df, **- litter_number**): remove "litter_number" and keep all of the other columns

* select(litters_df, **- starts_with("gd")**): remove everything that starts with "gd" and keep all of the other columns


## `select` and rename and `everything()`
```{r}
select(litters_df, group, litter_id = litter_number)

select(litters_df, group, litter_id = litter_number, everything())

rename(litters_df, litter_id = litter_number)

select(litters_df, gd0_weight, everything())

relocate(litters_df, litter_number)
```

* select(litters_df, group, **litter_id = litter_number**): keep group column and litter_number column, but rename 'litter_number' col to 'litter_id'

* select(litters_df, group, litter_id = litter_number, **everything()**): keep group col & keep litter_number col, but rename 'litter_number' col to 'litter_id' & keep everything that I have not mentioned yet. 
  * **rename(litters_df, litter_id = litter_number)** is also for renaming the variable and keeping everything else.

* select(litters_df, gd0_weight, everything()): **reorder the cols**. Give gd0_weight first, and then the other cols. When i need to put cols in a diff order but still *keep everything else*, i can use this code.
  * **relocate(litters_df, litter_number)** is also for bringing 'litter_number' first and keeping everything else.

## `select` vs `pull`

```{r}
select(litters_df, group)

pull(litters_df, group)
```

* **`select`**: get a *data frame* out and it doesn't care that there's only 1 col here.
* **`pull`**: would break the tidyverse logic by extracting a col from your data frame. and the output is just like a vector floating around -- it *doesn't exist inside of a data frame* anymore


# `filter`

Get rid of rows using `filter()`

```{r}
filter(litters_df, group == "Con7")

filter(litters_df, group != "Con7")

filter(litters_df, gd0_weight > 20)

filter(litters_df, group == "Con7" | group == "Con8")

filter(litters_df, group %in% c("Con7", "Con8"))

filter(litters_df, !(group == "Con7"))

filter(litters_df, group %in% c("Con7", "Con8"), gd0_weight > 20)

filter(litters_df, group %in% c("Con7", "Con8") & gd0_weight > 20)
```

* **filter(litters_df, group == "Con7")**: i want to keep inside everything here where group is equal to this specific value "Con7".
  * **`=`** means put this value into this variable VS **`==`** means check whether or not this thing is true/equal to. Filter would keep everything that this turns out to be true.
  * **`!=`** means check whether or not this is not equal to. So if run `filter(litters_df, group != "Con7")`, the output would be everything that group is not equal to Con7.

* We will often filter using comparison operators: **`<, <=, >, >=, ==, !=`**

* **`filter(litters_df, group == "Con7" | group == "Con8")`**: want to keep more than 1 col. **`|`** is saying 'if this is true OR if the next thing is true, we are gonna keep the result here.
  * **`filter(litters_df, group %in% c("Con7", "Con8"))`** is another way to do this. 
  
* **`filter(litters_df, !(group == "Con7"))`**: if want to negate the entire thing. It would give the opposite in the () when the thing in () is true. This is similar with and get the same output as using `group != "Con7"`

* **`filter(litters_df, group %in% c("Con7", "Con8"), gd0_weight > 20)`**: can use this to do multiple condition. 
  * Here means "AND", and when using "|" it means "OR". 
  * `filter(litters_df, group %in% c("Con7", "Con8") & gd0_weight > 20)` are the same with it.

## `drop_na` for drop missing rows

A very common filtering step requires you to omit missing observations. You can do this with filter, but I recommend using **drop_na** from the *tidyr package*:
```{r}
drop_na(litters_df)

drop_na(litters_df, gd0_weight)
```

* **`drop_na(litters_df)`** will remove any row with a missing value in data frame litters_df
* **`drop_na(litters_df, gd0_weight)`** will remove rows for which gd0_weight is missing.


# `mutate`

This is used to **add or change variable**.

```{r}
mutate(litters_df, 
       wt_gain = gd18_weight - gd0_weight, 
       group = str_to_lower(group))
```

* **`mutate(litters_df, wt_gain = gd18_weight - gd0_weight)`**
  * **`wt_gain = gd18_weight - gd0_weight`**: creating a new variable based on what we already have. New variables appear at the end of the dataset in the order that they are created. 
  * **`group = str_to_lower(group)`**: overwriting an old variable. Here, the new 'group' is to make the old 'group' into lower case of whatever the variable is.

* We can create a new variable and immediately refer to (or change) it.


# `arrange`

If my data is not organized in a helpful way, i can use `arrange`.

```{r}
arrange(litters_df, gd0_weight)

arrange(litters_df, group, gd0_weight)
```

* **`arrange(litters_df, gd0_weight)`**: reorganizing and ordered based on gd0_weight, from lowest to the next lowest...(ascending).
  * if want descending order: **`arrange(litters_df, desc(gd0_weight))`**
* **`arrange(litters_df, group, gd0_weight)`**: if we have nested data structures, this is helpful when we have groups. Here, first put 'group' in alphabetical order. Then, inside each group of 'group', put things in numeric order for gd0_weight.