---
title: "data importing"
author: "Shihui Peng"
date: "2023-09-27"
output: github_document
---

```{r, message=FALSE, echo=FALSE}
library(tidyverse)
```
The function we are going to use is in the readr package, which is loaded by default by the Tidyverse.

# Import the `FAS_litters.csv` csv using a relative path.

**read_csv():** first, tell 'read_csv' where to find this file, in (). If we put all datasets in this corresponding R project folder, we can use this relative path ("data/FAS_litters.csv")

```{r}
litters_df =
  read_csv("data/FAS_litters.csv")

litters_df = 
  janitor::clean_names(litters_df)
```

**litters_df = janitor::clean_names(litters_df):**
if the name has space, we can switch the name into snake case with this code.
Eg. Litter Number -> litter_number; Gd0 Weight -> gd0_weight; etc. 

# Import the same dataset using an absolute path.

```{r}
litter_df_abs = 
  read_csv("~/Desktop/data_wrangling_i/data/FAS_litters.csv")
```

This might work, but if we move this r project folder to another place, the relative path does not change (the direction is still from this project to this data file), but the absolute path would change. So litter_df is working. but litter_df_abs is not working in this way.

Therefore, **use relative path in a project** and **have all you need in the same place** to make things easy and your code reproducible.

## Learning assessment for importing dataset FAS_pups.csv
```{r learning_assessment_import_dataset}
pups_df = 
  read_csv("data/FAS_pups.csv")

pups_df =
  janitor::clean_names(pups_df)
```

# Look at data
```{r}
litters_df

head(litters_df)

tail(litters_df)
```

* **litters_df**: show 1st 10 rows by default, might not show not all of the columns
* **head(litters_df)**: show 1st six rows
* **tail(litters_df)**: show bottom six rows. more useful b/c when something goes wrong, something like weird missing or weird structure would often happen in the bottom rows

```{r, eval=FALSE}
view(litters_df)
```

If you want to look at the entire data set, use **view()**. A separate window will pull up and we can scroll through it.

# Look at a data summary

```{r}
str(litters_df)

skimr::skim(litters_df)
```

* **str(litters_df)**: looks at the structure of the data frame
* **skimr::skim(litters_df)**: useful for 1st time to look at a dataset. can give info about variable type, min, max, empty, n, mean, sd, p0/25/50/75/100, histograms

# Options in `read_*`

## `skip` and `col_names`

```{r}
litters_df =
  read_csv(
    "data/FAS_litters.csv",
    skip = 10,
    col_names = FALSE
  )
```

* **skip=**: read_csv assumes the 1st row in us data set is variable names and it starts reading things in from the very top row. If do *"skip=10"*, 1st 10 rows would be skipped. Now the 11th row is assumed as a collection of variable names by default. This would make troubles.
* **col_names = FALSE**: r would make up its own column names, such as x1, x2, .... But it still read in the last rows that we have here.
  * When you have a dataset which 1st 10 rows are variable explanations or data dictionary, we can skip these 10 rows and start importing us data. The 11th row would not be the variable name row but the 1st row of record we need.
  
## `na = c()`

```{r}
litters_df =
  read_csv(
    "data/FAS_litters.csv",
    na = c("NA", 99, " ", ".")
  )

litters_df
```

* **na = c(...)**: whenever you see NA, number value 99, space, or ., treat them as missing/NA

## `col_types`

```{r}
litters_df =
  read_csv(
    "data/FAS_litters.csv",
    col_types = 
      cols(
        'GD0 weight' = col_character(),Group = col_factor()
      )
  )

litters_df
```

* r would by default guess what kinds of data for each column, but it won't look through all rows. So if needed, we can change or define the column types for specific columns.
* **col_types = cols()** can be used:
  * **'GD0 weight' = col_character()**: since there is a space in (GD0 Weight), we need to use "" to include this variable name. This code is changing the column type of GD0 weight to character.
  * **Group = col_factor()**: changing the column type of Group to factor.
  
