---
title: "data importing"
author: "Shihui Peng"
date: "2023-09-27"
output: github_document
---

```{r, message=FALSE, echo=FALSE}
library(tidyverse)
```
The function we are going to use is in the readr package, which is loaded by default by the Tidyverse.

# Import the `FAS_litters.csv` csv using a relative path.

**read_csv():** first, tell 'read_csv' where to find this file, in (). If we put all datasets in this corresponding R project folder, we can use this relative path ("data/FAS_litters.csv")

```{r}
litters_df =
  read_csv("data/FAS_litters.csv")

litters_df = 
  janitor::clean_names(litters_df)
```

**litters_df = janitor::clean_names(litters_df):**
if the name has space, we can switch the name into snake case with this code.
Eg. Litter Number -> litter_number; Gd0 Weight -> gd0_weight; etc. 

# Import the same dataset using an absolute path.

```{r}
litter_df_abs = 
  read_csv("~/Desktop/data_wrangling_i/data/FAS_litters.csv")
```

This might work, but if we move this r project folder to another place, the relative path does not change (the direction is still from this project to this data file), but the absolute path would change. So litter_df is working. but litter_df_abs is not working in this way.

Therefore, **use relative path in a project** and **have all you need in the same place** to make things easy and your code reproducible.

## Learning assessment for importing dataset FAS_pups.csv
```{r learning_assessment_import_dataset}
pups_df = 
  read_csv("data/FAS_pups.csv")

pups_df =
  janitor::clean_names(pups_df)
```

# Look at data
```{r}
litters_df

head(litters_df)

tail(litters_df)
```

* **litters_df**: show 1st 10 rows by default, might not show not all of the columns
* **head(litters_df)**: show 1st six rows
* **tail(litters_df)**: show bottom six rows. more useful b/c when something goes wrong, something like weird missing or weird structure would often happen in the bottom rows

```{r, eval=FALSE}
view(litters_df)
```

If you want to look at the entire data set, use **view()**. A separate window will pull up and we can scroll through it.

# Look at a data summary

```{r}
str(litters_df)

skimr::skim(litters_df)
```

* **str(litters_df)**: looks at the structure of the data frame
* **skimr::skim(litters_df)**: useful for 1st time to look at a dataset. can give info about variable type, min, max, empty, n, mean, sd, p0/25/50/75/100, histograms

